---
title: "Comparison between Provisional and CFS 2012 Data"
author: "Parsons Brinckerhoff"
date: "October 14, 2015"
output: html_document
---

```{r setup}
library(knitr)
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
```

Version 4.0 of the Freight Analysis Framework --- based on the 2012 Commodity
Flow Survey and Economic Census --- was released last night. As of right now,
only the regional flows by commodity and mode for the base year are available.

There are many more things we need to have from the FAF besides these flows (for
instance, geographical definitions of the regions). However, for the first time
we can compare the "real" 2012 flows to what we have been including in our models.

** The results are not pretty.**

I downloaded the new FAF 4 dataset and aggregated the flows to the state level, to
account for possible changes in zonal definitions. I also applied two different
interpolation methodologies to the FAF 3.5:

  - `gdp` I used industry-level gross product information from the Bureau of Economic Analysis to extrapolate goods production from 2007 to 2012.
  - `lin` I linearly interpolated the 2007 values and 2015 forecasts to calculate a 2012 estimate.

```{r load_4, echo = FALSE}
faf4 <- read_csv(
  "../data_raw/faf4_data.csv", 
  col_types = list(
    fr_orig = col_character(),
    dms_orig = col_character(),
    dms_dest = col_character(),
    fr_dest = col_character(),
    fr_inmode = col_character(),
    fr_outmode = col_character(),
    sctg2 = col_character(),
    trade_type = col_numeric()
  )
) %>% 
  # only trucks
  # aggregate to state
  mutate(
    st_orig = substr(dms_orig, 1, 2),
    st_dest = substr(dms_dest, 1, 2),
    sctg = substr(sctg2, 1, 1)
  ) %>%
  filter(sctg != "9") %>%
  group_by(st_orig, st_dest, sctg) %>%
  summarise(
    tons_faf4 = sum(tons_2012, na.rm = TRUE),
    value_faf4 = sum(value_2012, na.rm = TRUE)
  ) %>%
  gather(var, flow, tons_faf4:value_faf4) %>%
  separate(var, c("type", "data"))
```

```{r load_35, echo = FALSE}
# read in historical gdp data
gdp <- readRDS("../data/gdp_output.rds") %>%
  filter(data_year == 2012) %>%
  rename(sctg2 = sctg)

# function to linearly interpolate
interpolate <- function(y0, y1, year){
  y <- y0 + (y1 - y0) * ( (year - 2007)/(2015 - 2007) )
  return(y)
}

faf35 <- read_csv(
  "../data_raw/faf35_data.csv", 
  col_types = list(
    fr_orig = col_character(),
    dms_orig = col_character(),
    dms_dest = col_character(),
    fr_dest = col_character(),
    fr_inmode = col_character(),
    fr_outmode = col_character(),
    sctg2 = col_character(),
    trade_type = col_numeric()
  )
) %>%
  select(dms_orig, dms_dest, sctg2, 
         value_2007, tons_2007, value_2015, tons_2015 ) %>%
  left_join(gdp) %>%

  # scale value and tons by gdp factor
  mutate(
    value_gdp = value_2007 * gdp,
    tons_gdp = tons_2007 * gdp,
    value_lin = interpolate(value_2007, value_2015, 2012),
    tons_lin = interpolate(tons_2007, tons_2015, 2012)
  ) %>%
  mutate(
    st_orig = substr(dms_orig, 1, 2),
    st_dest = substr(dms_dest, 1, 2),
    sctg = substr(sctg2, 1, 1)
  ) %>%
  filter(sctg != "9") %>%
  group_by(st_orig, st_dest, sctg) %>%
  summarise_each(funs(sum), value_gdp:tons_lin) %>%
  gather(var, flow, value_gdp:tons_lin) %>%
  separate(var, c("type", "data"))
```

In general, the two datasets are compatible with each other. Field names did not 
change, though the mode codes or the modes used seem to have changed. For this
reason, I look at freight moved by all modes. The table below shows twenty random 
state-to-state commodity flows from the combined dataset.

```{r joined, results = "asis"}
faf <- rbind_list(faf35, faf4) %>%
  spread(data, flow, fill = 0) 

kable(sample_n(faf, 10), caption = "Sample", digits = 2)
```


# Error Assessment

## Total Flows

The previous FAF 3.5 was based on the 2007 commodity flow survey and then
forecasted based on economic projections that did not include the recession from
2008-2009. Unsurprisingly, total tonnage of freight is substantially (around
20%) higher in the interpolated FAF 3.5 than in the FAF 4.0. However, the total 
value of this freight is only off by about 3.5%; this may in part be explained by
the near-zero inflation rates that have persisted through the recession.

```{r totals, results='asis'}
faf %>%
  group_by(type) %>%
  summarise_each(funs(sum), faf4:lin) %>%
  mutate(
    gdp_error = (gdp - faf4) / faf4 * 100,
    lin_error = (lin - faf4) / faf4 * 100
  ) %>%
  kable(caption = "Total FAF 2012 Flows", digits = 3)
```

If we look at the totals by 1-digit commodity code, then some patterns begin to 
emerge. Interpolating FAF 3.5 leads to a substantial overprediction in all
goods except for a mild underprediction in raw minerals (sctg 1...). However, the
total value of the goods shipped was fairly accurate in the interpolation with
two exceptions: an over-prediction in the value of manufactured goods (sctg 3...)
and a substantial underprediction in the value of raw minerals. All of these
observations can be explained by a loss of consumer demand (from the recession)
coupled with an energy boom.

```{r totals_sctg, results='asis'}
faf %>%
  group_by(type, sctg) %>%
  summarise_each(funs(sum), faf4:lin) %>%
  mutate(
    gdp_error = (gdp - faf4) / faf4 * 100,
    lin_error = (lin - faf4) / faf4 * 100
  ) %>%
  kable(caption = "Total FAF 2012 Flows by SCTG", digits = 3)
```

## State-to-State Flow Comparison

The error between FAF 4 and FAF 3.5 interpolations is not uniform for all
state-to-state pairs. The figure below shows a plot of the interpolations
for tons and value against the observed 2012 volume. It seems as though a major
source of error in the forecasts is optimism about flows that are near zero in
2012. Perhaps these were near zero in 2007 as well, but they were forecast for 
positive growth. Or alternatively, the recession caused small flows in 2007 to 
shut down completely. 


```{r error, fig.cap="Overall error in state-level flow prediction.", message=FALSE}
faf_p <- faf %>%
  gather(forecast, flow, gdp:lin) 

p <- ggplot(faf_p, aes(x = faf4 + 0.001, y = flow + 0.001)) +
  facet_grid(type ~ forecast) +
  scale_x_log10() + scale_y_log10() +
  xlab("FAF 4 Observed") + ylab("FAF 3.5 Interpolated") +
  theme_bw() +
  geom_abline(slope = 1, color = "grey")

p + geom_point(alpha = 0.2) + geom_smooth()
```

This is largely borne out in the commodity-level analysis below, except that the
problem is most pronounced with the raw minerals.


```{r error_sctg, fig.cap="Error in state-level flow prediction by SCTG."}
p + geom_point(aes(color = sctg), alpha = 0.2) + 
  geom_smooth(aes(color = sctg))
```

Put another way: FAF 3.5 interpolations of state-to-state flows seem adequate for 
large flows, but substantially over-predict small ones, especially flows relating 
to raw minerals. There are enough small flows to substantially affect the total
volume of trucks on the road.

# Lessons
We need to use the most updated FAF for all of our model work as soon as possible,
or else develop a new interpolation method.