---
title: "Zones"
author: "Parsons Brinckerhoff"
date: "October 15, 2015"
output: html_document
---
```{r setup}
library(dplyr)
library(tidyr)
library(foreign)
```



```{r faf_data}
load("../data/faf_data.Rdata")

prod <- FAF %>%
  group_by(dms_orig) %>%
  summarise(n = n(), value = sum(value), tons = sum(tons)) %>%
  mutate(
    state = substr(dms_orig, 1, 2)
  )
```

Edits to the FAF zone system built from the CFS regions:

  - Mobile to `012` from `013`
  - Alaska to `020` from `029`
  - Arkansas to `050` from `059`
  - Phoenix to `041` from `049`
  - Tuscon to `041` from `045`
  - San Diego to `063` from `069`
  - Fresno to `065` from `069`
  - Los Angeles to `061` from `062`
  - Sacramento to `062` from `064`
  - Denver to `081` from `082`
  - Hartford to `091` from `092`
  - New York (Connecticut) to `092` from `094`
  - Wilmington to `101` from `104`
  - District of Columbia to `111` from `115`
  - Jacksonville to `121` from `123`
  - Orlando to `123` from `124`
  - Tampa to `124` from `129`
  - Miami to `122` from `123`
  - Savannah to `132` from `134`
  - Idaho to `160` from `169`
  - St. Louis (Illinois) to `172` from `174`
  - Fort Wayne to `183` from `182`
  - Iowa to `190` from `199`
  - Kansas City (Kansas) to `201` from `203`
  - Wichita to `202` from `205`
  - Cincinatti (Kentucky) to `212` from `211`
  - Lexington to `211` from `213`
  - Lake Charles to `222` from `229`
  - Baton Rouge to `221` from `229`
  - New Orleans to `223` from `224`
  - Maine to `230` from `239`
  - Baltimore to `241` from `245`
  - Washington (Maryland) to `242` from `245`
  - Detroit to `261` from `262`
  - Minneapolis to `271` from `273`
  - Mississippi to `280` from `289`
  - Kansas City (Missouri) to `291` from `293`
  - St. Louis to `292` from `294`
  - Montana to `300` from `309`
  - Omaha to `311` from `314`
  - Las Vegas to `321` from `323`
  - Philadelphia (New Jersey) to `342` from `344`
  - New York (New Jersey) to `341` from `344`
  - New Mexico to `350` from `359`
  - Buffalo to `362` from `361`
  - New York to `363` from `364`
  - Raleigh to `373` from `374`
  - North Dakota to `380` from `389`
  - Cleveland to `392` from `391`
  - Dayton to `394` from `392`
  - Columbus to `393` from `391`
  - Oklahoma City to `401` from `404`
  - Tulsa to `402` from `405`
  - Portland to `411` from `414`
  - Pittsburgh to `422` from `424`
  - Philadelphia to `421` from `424`
  - New York (Pennsylvania) to `423` from `424`
  - Charleston to `451` from `459`
  - South Dakota to `460` from `469`
  - Memphis to `471` from `473`
  - Nashville to `472` from `474`
  - Austin to `481` from `489`
  - Port Arthur to `482` from `489`
  - Corpus Christi to `483` from `489`
  - Dallas to `484` from `482`
  - El Paso to `485` from `482`
  - Houston to `486` from `482`
  - Laredo to `487` from `489`
  - San Antonio to `488` from `489`
  - Salt Lake City to `491` from `494`
  - Vermont to `500` from `509`
  - Richmond to `501` from `515`
  - Virginia Beach to `512` from `515`
  - Washington (Virginia) to `513` from `515`
  - Seattle to `531` from `535` 
  - Portland (Washington) to `532` from `534`
  - West Virginia to `540` from `549`
  - Milwaukee to `551` from `553`
  - Wyoming to `560` from `569`
  
  
```{r read_adjusted}
cfs_zones <- read.dbf("../data_raw/shapefiles/faf4zone.dbf", as.is = TRUE)$F4Z
faf_zones <- names(table(FAF$dms_orig))


table(cfs_zones %in% faf_zones)
table(faf_zones %in% cfs_zones)
cfs_zones[!(cfs_zones %in% faf_zones)]
faf_zones[!(faf_zones %in% cfs_zones)]
```

  
  
  
  
  