---
title: "trucks_calculation"
author: "Gregory Macfarlane"
date: "November 24, 2015"
output: html_document
---

We follow the FAF 3 [traffic analysis methodology](http://faf.ornl.gov/fafweb/Data/Freight_Traffic_Analysis/chap3.htm)
published by ORNL.

```{r libs}
library(dplyr)
library(tidyr)
library(readr)
library(knitr)
```


# Factors
Read the truck allocation factors (Table 3-3) into memory, and reshape to make
it more useful. Also account for R's extremely annoying habit of converting
everything it can into a factor.
```{r allocation, results='asis'}
# DEFINE TRUCK ALLOCATION FACTORS -------------------------------------------
truck_allocation_factors <- read_csv("../data_raw/trucks/truckallocation.csv") %>%
  gather(vehicle_type, allocation_factor, -distance, convert = TRUE)
kable(truck_allocation_factors)
```


Read the truck equivalency factors found in Appendix A of the FAF3 Freight
Traffic Analysis report, which we'll need to convert from wide to tall format.
```{r equivalence}
truck_equivalency_factors <- read_csv("../data_raw/trucks/truckequivalence.csv") %>%
  gather(body_type, equiv_factor, auto:other, convert = TRUE)
kable(truck_equivalency_factors %>% sample_n(10))
```

Finally, read empty truck factors from Table 3-5 and define a helper function
to grab the appropriate ones by body type and flow direction. Again, convert
from the original wide to tall format on the fly.

```{r empty}
empty_truck_factors <- read_csv("../data_raw/trucks/emptytrucks.csv") %>%
  gather(vehicle_type, empty_factor, SU:TPT, convert = TRUE)
kable(empty_truck_factors %>% sample_n(10))
```


It will be much more efficient to join all of the factors to the data at one
time.
```{r factors}
truck_factors <- left_join(
  truck_allocation_factors, truck_equivalency_factors, 
  by = "vehicle_type"
) %>%
  left_join(empty_truck_factors, ., by = c("body_type", "vehicle_type"))
kable(truck_factors %>% sample_n(10))
```

# Application

We can walk through the process to convert a row of the FAF into loaded and
empty trucks. Recreate the conversion laid out in Table 3-6 of the FAF3 Freight Traffic
Analyis report.

```{r faf_row}
flow_records <- data_frame(
  id = 1, dms_orig=49, dms_dest=41, sctg="03", tons=1519.150, 
  value=1373.96, distance="(100,200]", crossing_type = "land_border",
  fr_inmode = NA, fr_outmode = NA
)
kable(flow_records)
```

Join the truck factors to the flow records table, `left_join` will expand
the LHS table as neccessary to match as many records as in needs to on the
RHS.
```{r expanded}
expandedtable <- left_join(
  flow_records, truck_factors, by = c("sctg", "distance", "crossing_type")
) 
kable(
  expandedtable %>%
    select(tons, distance, crossing_type, body_type, vehicle_type, empty_factor,
           allocation_factor, equiv_factor)
)
```

We now multiply the factors through the data.frame to get number of
trucks in each body type and each vehicle type. 
```{r truck_table}
truck_table <- expandedtable %>%
  mutate(
    # tons in FAF are given in kilotons
    trucks = 1000 * tons * allocation_factor * equiv_factor, 
    # As well as the number of empty trucks!
    empty_trucks = trucks * empty_factor
  )

kable(
  truck_table %>% 
    select(tons, body_type, vehicle_type, trucks, empty_trucks)
)
```

We care about vehicle type, but not body type. So let's group on the `id`
field as well as the `vehicle_type`, and then collapse everything else.
But we also want to keep the relevant fields from the original data.
because these are constant, we'll just take the first record in each group.
```{r outtable}
outtable <- truck_table %>%
  group_by(id, vehicle_type) %>%
  # we obviously want the sum of the trucks and the empty trucks.
  summarise(
    trucks = sum(trucks), empty_trucks = sum(empty_trucks), 
    dms_orig = dms_orig[1], dms_dest = dms_dest[1], sctg = sctg[1], 
    tons = tons[1], value = value[1]
  ) %>%
  gather(type, trucks, trucks:empty_trucks) %>%
  mutate(
    type = ifelse(
      type == "empty_trucks",  
      paste(vehicle_type, "empty", sep = ":"), 
      vehicle_type
    )
  )

kable(outtable)
```

It's difficult to calculate the average load per truck from this output, because
we would need to apply the same allocation factors that we used in creating it.
But grouping all trucks together, we get

```{r load_per_truck}
1519.15 * 1000 / sum(outtable$trucks[1:5]) # loaded trucks only
```

